A free/open-source Kazakh-Tatar machine translation system

### 2013-04-21 paper/paper.tex
add an example of a lexical selection rule
some small additions to new lrx stuff

For example (see figure~\ref{fig:bidix}), Kazakh құрал can mean an instrument, device, tool, or even weapon, all meanings corresponding to its Tatar cognate корал; however, it is also used in the compound ақпарат құралдары \eng{mass media} (literally, \eng{means of information}), which translates to Tatar as мәгълүмат чаралары (which has the same literal translation).  Hence, the Kazakh word құрал must have two entries in the bilingual lexicon: one that corresponds to Tatar корал and one that corresponds to Tatar чара.  A lexical selection rule that selects the translation чара when it occurs in a compound with ақпарат is written to ensure the correct translation.
%FIXME: Ilnar, we probably want an example of this rule

@@

>>>>

For example (see figure~\ref{fig:bidix}), Kazakh құрал can mean an instrument, device, tool, or even weapon, all meanings corresponding to its Tatar cognate корал; however, it is also used in the compound ақпарат құралдары \eng{mass media} (literally, \eng{means of information}), which translates to Tatar as мәгълүмат чаралары (which has the same literal translation).  Hence, the Kazakh word құрал must have two entries in the bilingual lexicon: one that corresponds to Tatar корал and one that corresponds to Tatar чара.  A lexical selection rule that selects the translation чара when it occurs in a compound with ақпарат is written to ensure the correct translation; this rule is shown in figure~\ref{fig:lrx}.

\begin{figure*}[htbp]
\hspace{3cm}\parbox[t]{0.7\textwidth}{{\tt
%\begin{texttt}
  <rule> \\
    <match lemma="ақпарат" tags="n.attr"/> \\
    <match lemma="құрал" tags="n.*.px3sp.*"><select lemma="чара"/></match> \\
  </rule>
}}%\end{texttt}
\caption{A lexical selection rule that selects чара as the translation of құрал if part of a compound with ақпарат.}
\label{fig:lrx}
%\vspace{-1em}
\end{figure*}

@@

The system currently has a total of 33 lexical selction rules.
